Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2018-05-16 revised_prop_letter_ajhs.tex
revised_prop_letter_ajhs.tex Published version except for the policy issue...

\begin{table}
\caption{\label{tab:probability-measures}Possible probability measures on
a Hilbert space of dimension~$d=3$, where $\bar{\mu}_{2}'$ and
$\bar{\mu}_{3}$ are QIVPMs while $\bar{\mu}_{0}$, $\bar{\mu}_{1}$,
and $\bar{\mu}_{2}$ are not. Events are listed in the column labeled
by $P$.}
\centering{}%
\begin{tabular}{cccccc}
\toprule 
\addlinespace
$P$  & $\bar{\mu}_{0}\left(P\right)$  & $\bar{\mu}_{1}\left(P\right)$  & $\bar{\mu}_{2}\left(P\right)$  & $\bar{\mu}_{2}'\left(P\right)$  & $\bar{\mu}_{3}\left(P\right)$\tabularnewline
\midrule
\midrule 
\addlinespace
$\mathbb{0}$  & $\imposs$  & $\imposs$  & $\imposs$  & $\imposs$  & $\imposs$\tabularnewline
\midrule 
\addlinespace
\begin{tabular}{c}
All one-dimensional\tabularnewline
projectors\tabularnewline
\end{tabular} & $\left[0,0\right]$  & $\left[0,\tfrac{1}{4}\right]$  & $\left[0,\tfrac{1}{3}\right]$  & $\left[\tfrac{1}{3},\tfrac{1}{3}\right]$  & $\left[0,\tfrac{1}{2}\right]$\tabularnewline
\midrule 
\addlinespace
\begin{tabular}{c}
All two-dimensional\tabularnewline
projectors\tabularnewline
\end{tabular} & $\left[1,1\right]$  & $\left[\tfrac{3}{4},1\right]$  & $\left[\tfrac{2}{3},1\right]$  & $\left[\tfrac{2}{3},\tfrac{2}{3}\right]$  & $\left[\tfrac{1}{2},1\right]$\tabularnewline
\midrule 
\addlinespace
$\mathbb{1}$  & $\necess$  & $\necess$  & $\necess$  & $\necess$  & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
The bound $\delta<\frac{1}{3}$ is tight as it is possible to construct
a $\frac{1}{3}$-deterministic QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$.
For example, $\bar{\mu}_{2}'$ defined in Table~\ref{tab:probability-measures}
is a valid $\frac{1}{3}$-deterministic QIVPM. When $\delta\geq\frac{1}{3}$,
i.e, when the uncertainty in measurements becomes so large, it becomes
possible to map every observable to some (quite inaccurate) probability
interval, thus invalidating the Kochen-Specker theorem. We can summarize
and illustrate the above arguments using Fig.~\ref{fignoname}.

@@

As is the case for conventional, infinitely-precise, quantum
probability measures, the theorem is only applicable to dimensions
$d \geq 3$. Indeed when the Hilbert space has dimension 2, it is
straightforward to construct a 0-deterministic QIVPM as
follows. Consider a non-contextual hidden variable model for $d=2$
(e.g., as proposed by Bell or
Kochen-Specker~\cite{BELL_1966,kochenspecker1967}). Such a
two-dimensional model assigns definite values to all observables at
all times, and hence assigns a \emph{determinate} probability (0 or 1)
to each event. This probability measure directly induces a
0-deterministic QIVPM by changing 0 to $\imposs$ and 1 to
$\necess$.  It follows that every 0-deterministic QIVPM is
$\delta$-deterministic.
%% ajh: Fixed garbled sentence.

@@

\begin{thm}[Range of Ultramodular Functions]\label{thm:convex-uncountable}For
any ultramodular function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$,
either $\mathscr{I}=\mathscr{I}_{0}$ as defined in Eq.~(\ref{eq:3-value-intervals}) or
$\mathscr{I}$ contains uncountably many intervals.\end{thm}

>>>>

\begin{table}
\caption{\label{tab:probability-measures}Possible probability measures on
a Hilbert space of dimension~$d=3$, where $\bar{\mu}_{2}'$ and
$\bar{\mu}_{3}$ are QIVPMs while $\bar{\mu}_{0}$, $\bar{\mu}_{1}$,
and $\bar{\mu}_{2}$ are not. Events are listed in the column labeled
by $P$.}
\centering{}%
\begin{tabular}{cccccc}
\toprule 
\addlinespace
$P$  & $\bar{\mu}_{0}\left(P\right)$  & $\bar{\mu}_{1}\left(P\right)$  & $\bar{\mu}_{2}\left(P\right)$  & $\bar{\mu}_{2}'\left(P\right)$  & $\bar{\mu}_{3}\left(P\right)$\tabularnewline
\midrule
\midrule 
\addlinespace
$\mathbb{0}$  & $\imposs$  & $\imposs$  & $\imposs$  & $\imposs$  & $\imposs$\tabularnewline
\midrule 
\addlinespace
\begin{tabular}{c}
All one-dimensional\tabularnewline
projectors\tabularnewline
\end{tabular} & $\left[0,0\right]$  & $\left[0,\tfrac{1}{4}\right]$  & $\left[0,\tfrac{1}{3}\right]$  & $\left[\tfrac{1}{3},\tfrac{1}{3}\right]$  & $\left[0,\tfrac{1}{2}\right]$\tabularnewline
\midrule 
\addlinespace
\begin{tabular}{c}
All two-dimensional\tabularnewline
projectors\tabularnewline
\end{tabular} & $\left[1,1\right]$  & $\left[\tfrac{3}{4},1\right]$  & $\left[\tfrac{2}{3},1\right]$  & $\left[\tfrac{2}{3},\tfrac{2}{3}\right]$  & $\left[\tfrac{1}{2},1\right]$\tabularnewline
\midrule 
\addlinespace
$\mathbb{1}$  & $\necess$  & $\necess$  & $\necess$  & $\necess$  & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
The bound $\delta<\frac{1}{3}$ is tight as it is possible to construct
a $\frac{1}{3}$-deterministic QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$.
For example, $\bar{\mu}_{2}'$ defined in Table~\ref{tab:probability-measures}
is a valid $\frac{1}{3}$-deterministic QIVPM. When $\delta\geq\frac{1}{3}$,
i.e, when the uncertainty in measurements becomes too large, it becomes
possible to map every observable to some (quite inaccurate) probability
interval, thus invalidating the Kochen-Specker theorem. We can summarize
and illustrate the above arguments using Fig.~\ref{fignoname}.

@@

As in the case for conventional, infinitely-precise, quantum
probability measures, the theorem is only applicable to dimensions
$d \geq 3$. Indeed when the Hilbert space has dimension 2, it is
straightforward to construct a 0-deterministic QIVPM as
follows. Consider a non-contextual hidden variable model for $d=2$
(e.g., as proposed by Bell or
Kochen-Specker~\cite{BELL_1966,kochenspecker1967}). Such a
two-dimensional model assigns definite values to all observables at
all times, and hence assigns a \emph{determinate} probability (0 or 1)
to each event. This probability measure directly induces a
0-deterministic QIVPM by changing 0 to $\imposs$ and 1 to
$\necess$.  It follows that every 0-deterministic QIVPM is
$\delta$-deterministic.
%% ajh: Fixed garbled sentence.

@@

\begin{thm}[Range of Ultramodular Functions]\label{thm:convex-uncountable}For
any ultramodular function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$,
either $\mathscr{I}=\mathscr{I}_{0}$ as defined in Eq.~(\ref{eq:3-value-intervals}), or
$\mathscr{I}$ contains uncountably many intervals.\end{thm}
